---
author: "regmods-030 - justin harding"
date: "26 July 2015"
output: pdf_document
---
```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F, results='hide'}
# make this an external chunk that can be included in any file
library(knitr)
library(car)
```

```{r echo=FALSE}
fit.wt.hp.am <- lm(mpg ~ wt + hp + factor(am) + factor(am)*wt, mtcars)
```

## Executive Summary
From the supplied data set of a collection of cars, we are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). We attempt to answer the following two questions:

* Is an automatic or manual transmission better for MPG
* Quantify the MPG difference between automatic and manual transmissions

The conclusion reached from the data given is that there is no significant statistical difference between fuel consumption of manual and automatic transmission for the average vehicle.

However further investigation shows that automatic transmission gives better fuel consumption cars that weigh less/have less horspower and manual transmission is better for cars that weigh more/have more power, although there is a very wide variation between min and max mpg values in our models. Summarised in the following table.

weight|hp|trans|min mpg (conf interval 95%)|max mpg (conf interval 95%)
------|--|-----|---------------------------|---------------------------
2000lb|70|man|`r round(fit.wt.hp.am$coef[1] + fit.wt.hp.am$coef[2] * 2 + fit.wt.hp.am$coef[3] * 70 - 2 * summary(fit.wt.hp.am)$sigma, 2)`|`r round(fit.wt.hp.am$coef[1] + fit.wt.hp.am$coef[2] * 2 + fit.wt.hp.am$coef[3] * 70 + 2 * summary(fit.wt.hp.am)$sigma, 2)`
2000lb|70|auto|`r round(fit.wt.hp.am$coef[1] + fit.wt.hp.am$coef[4] + (fit.wt.hp.am$coef[2] + fit.wt.hp.am$coef[5]) * 2 + fit.wt.hp.am$coef[3] * 70 - 2 * summary(fit.wt.hp.am)$sigma, 2)`|`r round(fit.wt.hp.am$coef[1] + fit.wt.hp.am$coef[4] + (fit.wt.hp.am$coef[2] + fit.wt.hp.am$coef[5]) * 2 + fit.wt.hp.am$coef[3] * 70 + 2 * summary(fit.wt.hp.am)$sigma, 2)`
4000lb|140|man|`r round(fit.wt.hp.am$coef[1] + fit.wt.hp.am$coef[2] * 4 + fit.wt.hp.am$coef[3] * 140 - 2 * summary(fit.wt.hp.am)$sigma, 2)`|`r round(fit.wt.hp.am$coef[1] + fit.wt.hp.am$coef[2] * 4 + fit.wt.hp.am$coef[3] * 140 + 2 * summary(fit.wt.hp.am)$sigma, 2)`
4000lb|140|auto|`r round(fit.wt.hp.am$coef[1] + fit.wt.hp.am$coef[4] + (fit.wt.hp.am$coef[2] + fit.wt.hp.am$coef[5]) * 4 + fit.wt.hp.am$coef[3] * 140 - 2 * summary(fit.wt.hp.am)$sigma, 2)`|`r round(fit.wt.hp.am$coef[1] + fit.wt.hp.am$coef[4] + (fit.wt.hp.am$coef[2] + fit.wt.hp.am$coef[5]) * 4 + fit.wt.hp.am$coef[3] * 140 + 2 * summary(fit.wt.hp.am)$sigma, 2)`


## Procedure
The Rmd file used to generate this document can be found at https://github.com/justinharding/regmods-030_project.git
I have left out large amounts of information I used for this project to pare down to 5 pages. I hope what I have left still makes sense!

Our null hypothesis H~0~ : mpg(man) = mpg(auto)
and our alternate hypothesis H~a~ : mpg(man) != mpg(auto)


First check the data using `str(mtcars)`
```{r, echo=FALSE}
str(mtcars)
```

Then we create a fit to test all of the adjusted models using `lm(mpg ~ ., mtcars)`.
```{r, echo = FALSE}
fit <- lm(mpg ~ ., mtcars)
```

From the scatterplots generated using `pairs(mtcars)` (output in appendix), it appears the predictors most correlated wth _mpg_ (miles per gallon) are _wt_ (weight), _cyl_ (no. cylinders), _disp_ (displacement), and _hp_ (horsepower). In the [adjusted fits] in the appendix from this model `lm(mpg ~ ., mtcars)`, only _wt_ was even slightly correlated after effects of all other fits were taken out of predictor and outcome. We will try the models in the table below. Summaries from models can be found in the appendix.

fit.wt|fit.wt.hp|fit.wt.hp.cyl|fit.wt.hp.disp
------|---------|-------------|--------------
`mpg ~ wt`|`mpg ~ wt + hp`|`mpg ~ wt + hp + cyl`|`mpg ~ wt + hp + disp`

```{r echo = FALSE}
fit.wt <- lm(mpg ~ wt, mtcars)
fit.wt.hp <- update(fit, mpg ~ wt + hp)
fit.wt.hp.cyl <- update(fit, mpg ~ wt + hp + cyl)
fit.wt.hp.disp <- update(fit, mpg ~ wt + hp + disp)
```

From the output in the appendix both _cyl_ and _disp_ do not add anything useful to the model. An excerpt from the correlation matrix shows us that they are both quite highly correlated with existing terms.
```{r, echo = FALSE}
cor(mtcars)[c(4,6),c(1:6)]
```
Output from [anova] in the appendix shows us also that the extras terms do not help. Output from vif follows. I have concatenated the outputs to save space.

`sqrt(vif(fit.wt.hp.cyl))`
`sqrt(vif(fit.wt.hp.disp))`
`sqrt(vif(fit.wt.hp))`
```{r, echo = FALSE}
c(sqrt(vif(fit.wt.hp.cyl)), sqrt(vif(fit.wt.hp.disp)), sqrt(vif(fit.wt.hp)))
```

Our best model appears to be `fit.wt.hp <- lm(mpg ~ wt + hp, mtcars)`.

The [residuals] plot in the appendix shows that the residuals appear to be uncorrelated.  The quantile plot shows that there are issues with the data. Fiat 128, Toyota Corolla and Chrysler Imperial are all quite a way off the normal. This can also be seen from the dfbetas output. The hatvalues plot shows us that the Maserati Bora is well off the normal.

Now consider the effect of transmission type. See [transmission summary] in the appendix. 
```{r, echo = FALSE}
fit.wt.hp.am.same.slope <- lm(mpg ~ wt + hp + factor(am), mtcars)
```

## Conclusion

From the model `lm(mpg ~ wt + hp + factor(am))` for _factor(am)1_, we have confidence interval of `r round(confint(fit.wt.hp.am.same.slope)[4,1], 3)` to  `r round(confint(fit.wt.hp.am.same.slope)[4,2], 3)` and we have a p value greater than 0.05 which means we **cannot** rule out the null hypothesis. This in turn means we cannot rule out the possibility that transmission type has no effect on fuel consumption.


```{r echo=FALSE}
fit.wt.hp.am <- lm(mpg ~ wt + hp + factor(am) + factor(am)*wt, mtcars)
```

However if we allow our models to have different slopes, manual cars and automatic cars do have different characteristics. Manual cars have fuel consumption of `r fit.wt.hp.am$coef[1]` `+` `r fit.wt.hp.am$coef[2]` `* wt +` `r fit.wt.hp.am$coef[3]` `* hp` and automatic cars have fuel consumption of `r fit.wt.hp.am$coef[1] + fit.wt.hp.am$coef[4]` `+ (` `r fit.wt.hp.am$coef[2] + fit.wt.hp.am$coef[5]` `) * wt + ` `r fit.wt.hp.am$coef[3]` ` * hp`

## Appendix
Note: I assume that tables can be considered as figures for inclusion in Appendix.

**Scatter Plots**
```{r, echo=FALSE}
pairs(mtcars)
```

**Correlation Matrix**
```{r}
cor(mtcars)
```

**Coefficients for model `mpg ~ .`**
```{r, echo=FALSE}
kable(summary(fit)$coef, booktabs = TRUE)
```

**Coefficients for models `mpg ~ wt, mpg ~ wt + hp, mpg ~ wt + hp + disp, mpg ~ wt + hp + cyl`**
```{r, echo = FALSE}
kable(summary(fit.wt)$coef, booktabs = TRUE)
kable(summary(fit.wt.hp)$coef, booktabs = TRUE)
kable(summary(fit.wt.hp.disp)$coef, booktabs = TRUE)
kable(summary(fit.wt.hp.cyl)$coef, booktabs = TRUE)
```

**Anovas `anova(fit.wt, fit.wt.hp, fit.wt.hp.cyl)` and `anova(fit.wt, fit.wt.hp, fit.wt.hp.disp)`**
```{r, echo = FALSE}
anova(fit.wt, fit.wt.hp, fit.wt.hp.cyl)[1:5]
anova(fit.wt, fit.wt.hp, fit.wt.hp.disp)[1:5]
```

**Residuals Plots**
```{r, echo=FALSE}
par(cex = 0.5)
par(mfrow=c(2,2))
plot(resid(fit.wt.hp))
plot(density(resid(fit.wt.hp)))
```

**Excerpt from dfbetas and the hatvalues plot**
```{r, echo=FALSE}
round(dfbetas(fit.wt.hp)[15:23,], 3)
par(mfrow=c(2,2))
plot(fit.wt.hp, which = 2)
plot(hatvalues(fit.wt.hp))
```

**Models including transmission**
mpg ~ wt + hp + factor(am) and mpg ~ wt + hp + factor(am) + factor(am)*wt
```{r, echo = FALSE}
summary(fit.wt.hp.am.same.slope)$coef
summary(fit.wt.hp.am)$coef
```
